\documentclass[12pt]{article}



\usepackage[T1]{fontenc}
\usepackage{amsfonts, amsmath, amssymb}
\usepackage{multirow}
\usepackage{epsfig}
\usepackage{subfigure}
\usepackage{graphicx}
\usepackage{hyperref}
\usepackage{parskip}
\usepackage{booktabs}
\usepackage{longtable}
\usepackage[utf8]{inputenc}
\usepackage[english]{babel}
% \usepackage[document]{ragged2e}
\usepackage{verbatim, rotating, paralist}
\usepackage{enumerate}

\usepackage{enumerate}


\usepackage{pdfsync}
\usepackage{latexsym}
\usepackage{amsthm}
\usepackage{mathabx}

\usepackage{stmaryrd}
\usepackage{mathrsfs}
\usepackage{dsfont}
\usepackage{fancyhdr}
\usepackage{color}

\usepackage{parskip}
\usepackage{anysize, indentfirst, setspace}
\usepackage[right=1.75cm, left=1.75cm, top=3cm, bottom=3cm]{geometry}
\usepackage{appendix}

\usepackage{enumitem}
\setlist{nosep}

\renewcommand{\topfraction}{.85}
\renewcommand{\bottomfraction}{.7}
\renewcommand{\textfraction}{.15}
\renewcommand{\floatpagefraction}{.66}
\renewcommand{\dbltopfraction}{.66}
\renewcommand{\dblfloatpagefraction}{.66}




% \pagestyle{fancyplain}
% \rhead{\hfill \small \emph{MIDS NUMBER -- Fall 2019}}
\cfoot{}

% \renewcommand{\headrulewidth}{0pt}



%-------------------------- BEGIN DOCUMENT ----------------------------------%
\begin{document}


\singlespacing






%------------------------- HEADER ---------------------------------%
\thispagestyle{empty}
\begin{minipage}[t]{.5\textwidth}
	Adriane Fresh \\
	 Assistant Research Professor\\
	 Office Hours: TBD\\
     \vspace*{0.1cm}
\end{minipage}
\begin{minipage}[t]{.5\textwidth}
	\begin{flushright}  IDS 541\\
Fall 2025\\
    \vspace*{0.1cm}
\end{flushright}
\end{minipage}


% line
\line(1,0){499}

\vspace{.35in}

\begin{center}
	\textbf{\LARGE{Practical Data Science II} }
\end{center}





%--------------------------------------------COURSE DESCRIPTION--------------------------------------------------%

\section{Course Description}

Practical Data Science II is a flipped-classroom, exercise and project-focused course. Building on the computational thinking skills developed in Practical Data Science I, this course introduces students to a range of computational inquiry methods, including network analysis, geospatial analysis, and natural language processing (NLP). Throughout, the focus will be on developing hands-on experience implementing these methods with messy real-world data to ensure students are prepared to deploy these tools to answer the questions they care about. Requirements: Practical Data Science I, introductory statistics course.

\section{Class Schedule}

Please visit \href{http://www.practicaldatascience.org/ids591_specific/class_schedule_591.html}{https://www.practicaldatascience.org/ids590\_specific/class\_schedule\_591.html} for detailed class schedule.

\section{Class Organization}

Data science is an applied discipline, and so this will be an intensely applied class with \emph{lots} of hands-on exercises.

To make it possible for us to work through problems together as they arise, we will dedicate most of our class time to completing these exercises in small groups. That means that students will be required to read instructional material \emph{before every class} so they will be ready to do these exercises. This is what is referred to as ``flipping the classroom.''

In order to make this class organization work, it will be \textbf{\emph{critically}} important that students do their assigned readings before \emph{every} class, and as discussed below, this will be reflected in how grades are assigned in this class. Students who do not complete their assigned readings and tutorials before each class should not expect to receive good grades, regardless of performance on project assignments.

This class is organized around having two (synchronous) class sessions every week. While the plan is for most of these will be in person, some classes will inevitably end up needing to be held online. \textbf{Synchronous attendance, whether classes online or in person, is required unless you are unable to participate synchronously due to extenuating circumstances (such as an internet connection that will not support synchronous participation (for online classes) or illness (for in person classes)).} 

With that said, everyone's health and safety is of course our first priority, so while it is very important you attend class whenever possible, you should \textbf{never} hesitate to stay home if you're not feeling well. If you are not feeling well and need to miss class -- or need to miss class for covid related reasons (e.g. quarantine) -- please reach out to me \textbf{before class} so that we can make a plan to make sure you're fully supported!


%--------------------------------------------COURSE ASSIGNMENTS------------------------------------------------%
\section{Assignments \& Grading}

\subsection{Participation (20\% of Grade)}

Note that a major component of good participation is good \emph{preparation}. Because we will mostly reserve class time for hands-on exercises, it is absolutely critical that students do their assigned readings before \emph{every} class. Students who do not work through the instructional materials they have been assigned before class will not only get very little out of the hands-on exercises designed to reinforce the assigned materials, but they will also undermine the learning of the students they are asked to work with. With that in mind, students who do not complete their assigned readings before every class should expect to see this reflected in their participation grades.

Participation will be graded as follows:\footnote{This rubric is adapted from that of Duke Political Science Professor Adriane Fresh.}

\textbf{A range.}  You are fully \emph{and consistently} engaged in class discussion and exercises.  You both listen and contribute actively.  You are well-prepared for class.  Having done more than merely read the material, you have spent time thinking \emph{carefully and deeply} about the material's relationship to other materials and ideas presented in previous classes. You are not only able to answer questions about the material, but also come to class with thoughtful questions.  When working in teams, you work \emph{with} your partner. If your partner is struggling with an exercise, you help them understand the material rather than just completing the material on your own. If you are struggling with material, you ask for help (both from the instructor and your fellow students) and do not simply lean on your partner to complete the exercise. \\

\textbf{B range.}  You are engaged in class discussion and exercises.  You listen and contribute regularly.  You come well-prepared to class having read the material and your contributions show your familiarity, but your level of engagement lacks the depth accumulated through extra time spent thinking about the material.  When working in teams, you work \emph{with} your partner when they have a similar level of understanding, but do not always invest in helping a struggling partner to understand the material. You often ask for help when you are struggling, but other times you let your partner just complete the exercise. \\

\textbf{C range.}  You have met the minimum requirements of participation.  You are usually, but not always prepared.  You participate sometimes, but not regularly.  The comments that you offer show a basic familiarity with the materials, but do not help to build a coherent or productive discussion.  When working in teams, you only sometimes work \emph{with} your partner. When your partner is struggling, you often just do the exercise yourself. If you are struggling, you often do not ask for help and allow your partner to take over the exercise. \\

\textbf{D range.}  You have not met the minimum requirements of participation.  You are unprepared for class.  You have not read with the material with sufficient engagement to know even the most basic elements.  When working in teams, you do not attempt to work \emph{with} your partner. When your partner is struggling, you just do the exercise yourself. If you are struggling, you do not ask for help and allow your partner to take over the exercise.\\

\textbf{As should be clear from this rubric, above all it is important to emphasize that participation is evaluated on the basis of \emph{quality} and \emph{consistently}, \emph{not} quantity. Moreover, when completing in-class exercises, good participation is not about finishing first or without ever asking for help; good participation in in-class exercises is about helping your partner understand the material, and asking for help when you need it.}

\subsection{Quizzes (20\%)}

To ensure students are doing their readings in advance of class, from time to time we will start class with short quizzes. These quizzes are designed to be relatively straightforward if you did the readings—they won't be full of gotcha questions—but will require you to have done the readings.

\subsection{Exercises (35\% of Grade)}

Over the course of the semester, students will be asked to complete a number of small assignments as homework. These assignments will, in total, be worth 35\% of student grades. \textbf{Note:} because of the way students get autograder feedback before final submission and because the lowest exercise grade gets dropped, grades on exercises tend to converge towards 100\%. In light of that, I will generally adjust exercise scores down 5 percentage points when doing grade calculations so that getting all available points on an exercise (reported as 100\% on gradescope) maps to a 95\% solid A (but not a 100\%).

\subsection{Team Data Science Project  (25\% of Grade)}

Around mid-semester, students will be assigned a large team Data Science Project. The goal and general framework for this team project will be provided to students, but the project will require students to complete the analysis component of a full data science project, including gathering data, cleaning and merging that data, analyzing the data, and presenting results.

\subsection{Late Assignments, Make Up Exams and Extra Credit}


\textbf{Grading}

All assignments will be given a numerical score on a 0-1 scale.  These scores will be multiplied by the value of the assignment (see above) and the following scale will be used to assign a final letter grade.  \\


\textbf{Late Assignment}

All late assignments will be penalized 10\% per day the assignment is late, up to a maximum penalty of 50\%.

The final deadline for accepting assignments that are more than one week late is at the discretion of the instructor and may vary by assignment. 

Exceptions to these late penalties may be made for students dealing with exceptional circumstances (illness for themselves or family, etc.) — if you are dealing with a difficult situation, please feel free to contact me to discuss your situation.

\textbf{Dropping Lowest Scores}

To accommodate the fact that life happens, at the end of the semester, I will drop each student's lowest Quiz \emph{and} each student's lowest Exercise \textbf{that have been completed.} Essentially, this is a free pass for one quiz you miss or whiff, and a free pass on one exercise even if you submit it very late. 

But it is \emph{not} a free pass to \emph{skip} an Exercise — uncompleted Exercises are not eligible for being dropped.

\textbf{Quizzes and Absentees}

As detailed above, attendance in class is always required barring injury, illness, or other significant conflict. 

To ensure fairness: 

\begin{itemize}
	\item students who are not present in class will not be allowed to make up quizzes
	\item except if the student has reached out to me \textbf{before} class to alert me to their absence and the reason for it. If the reason is it deemed to be grounds for an excused absence, the student will be able to take the quiz at a different time.
	\item students who are not physically in class are also \textbf{NOT} allowed to take quizzes remotely without express permission. If you are found to have taken a quiz when not in class, not only will you receive a zero on that quiz, but it may be considered an honor code violation. Note that because classes are recorded (for students with excused absences) and gradescope tracks IP addresses, this is not a hard thing to figure out.
\end{itemize}


\section{Texts}

This year we will be working exclusively from \emph{Practical Data Science with Python}, a textbook being developed by myself and Kyle Bradbury. So you don't have to buy anything!

We will also do some readings from \href{https://www.amazon.com/Code-Language-Computer-Hardware-Software/dp/0735611319}{Code: The Hidden Language of Computer Hardware and Software} by Petzold, Charles. It's a fun book and not very expensive, but we won't use it a lot so copies of relevant chapters will be provided if you don't want to buy it.

\section{Course Schedule}

Because one aim of this course is to ensure that all MIDS students have a solid foundation for their time at Duke, the exact organization of this course is likely to change regularly as the course proceeds. Students will therefore be expected to regularly (i.e. before every class) check on the updated course schedule (which will include assignments for the next class) at \href{https://www.practicaldatascience.org}{www.practicaldatascience.org}.

\section{Honor Policy}

Duke University is a community dedicated to scholarship, leadership, and service and to the principles of honesty, fairness, respect, and accountability. Citizens of this community commit to reflect upon and uphold these principles in all academic and nonacademic endeavors, and to protect and promote a culture of integrity.

Remember the \href{https://studentaffairs.duke.edu/conduct/about-us/duke-community-standard}{Duke Community Standard} that you have agreed to abide by:

\begin{itemize}
	\item I will not lie, cheat, or steal in my academic endeavors;
	\item I will conduct myself honorably in all my endeavors; and
	\item I will act if the Standard is compromised.
\end{itemize}

Cheating on exams or plagiarism on homework assignments, lying about an illness or absence and other forms of academic dishonesty are a breach of trust with classmates and faculty, violate the Duke Community Standard, and will not be tolerated. Such incidences will result in a 0 grade for all parties involved. Additionally, there may be penalties to your final class grade along with being reported to the MIDS program directors.

\section{Disability Statement}

In an effort to prevent students with disabilities from having to explain and justify their condition separately to each of their various instructors, Duke has centralized disability management in the \href{https://access.duke.edu/students}{Student Disabilities Access Office}. If you think there is a possibility you may need an accommodation during this course, please reach out to their office as soon as possible (processing can take a little time).

Medical information shared with the SDAO are strictly confidential, and if SDAO determines an accommodation is appropriate, faculty members will simply be informed of the accommodation they are required to provide, not the underlying medical reason for the accommodation.

If you have any problems with SDAO, please let me know as soon as possible.

\section{Student Signature}

I, the undersigned, confirm I have read and understand the expectations of this class.
\\
\vspace{1cm}

Name: \_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_
\\
\vspace{1cm}

Signature: \_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_
\\
\vspace{1cm}

Date: \_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_

\vspace{2cm}


I, the undersigned, confirm I have also read and understand the \href{https://nickeubank.github.io/practicaldatascience_book/notebooks/PDS_not_yet_in_coursera/99_advice/llms.html}{chatGPT and You} reading.
\\
\vspace{1cm}

Name: \_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_
\\
\vspace{1cm}

Signature: \_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_
\\
\vspace{1cm}

Date: \_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_


% \section{Learning Goals}
%
% In Data Science today, the only constant is change. With that in mind, in this course we will not only learn \emph{how} popular tools work, but also:
% \begin{itemize}
% 	\item the logic that underlies their operation (so when new situations arise you will have a \emph{generalized} understanding of the tool you can use to reason through your problem), and
% 	\item how to find help on your own.
% \end{itemize}
%
% In particular, by the end of this course, you will have developed the following abilities in each topic area:
%
% \textbf{The Command Line}\\
% \emph{Main Takeaway: The Command Line is just a way to interact with your operating system with text instead of with a mouse.}
% \begin{itemize}
% 	\item Explain the value of the command line
% 	\item Manipulate files and work with command-line-only tools
% 	\item Anticipate the likely syntax of new tools you may come across
% \end{itemize}
%
% \textbf{\texttt{numpy}}\\
% \emph{Main Takeaway: numpy is what makes Python useable for data science.}
% \begin{itemize}
% 	\item Explain \emph{why} numpy and pandas are so crucial to data science in Python
% 	\item Manipulate vectors and matrices with \texttt{numpy}
% \end{itemize}
%
% \textbf{\texttt{pandas}}\\
% \emph{Main Takeaway: pandas is a hack, so if it drives you nuts, it's not your fault.}
% \begin{itemize}
% 	\item Read in data of various formats with \texttt{pandas}
% 	\item Clean, organize, and reshape real-world data with \texttt{pandas}
% 	\item Move back and forth from \texttt{pandas} to \texttt{numpy}
% 	\item Pass data from \texttt{pandas} and \texttt{numpy} to \texttt{scikit-learn} functions
% \end{itemize}
%
% \textbf{Git and Github}\\
% \emph{Main Takeaway: Bundling changes into discrete chunks is incredibly powerful}
% \begin{itemize}
% 	\item Not sure yet...
% \end{itemize}
%
%
% \textbf{Getting Help Online}\\
% \emph{Main Takeaway: Asking for help effectively takes effort}
% \begin{itemize}
% 	\item Find appropriate forums for different types of questions
% 	\item Compose requests for help that are likely to get useful responses using Minimal Working Examples (MWEs) and proofs of effort.
% \end{itemize}
%
% \textbf{Debugging}\\
% \emph{Main Takeaway: Debugging as an \textbf{active} exercise}
% \begin{itemize}
% 	\item Isolate and analyze bugs quickly
% \end{itemize}
%
% \textbf{Workflow Management}\\
% \emph{Main Takeaway: Projects change, so a good workflow must be adaptive}
% \begin{itemize}
% 	\item Organize data science projects in a manner that is robust to future changes
% 	\item Organize, document, and comment projects to allow others (and your future self) to easily understand project organization
% \end{itemize}
%
% \textbf{Defensive Programming}\\
% \emph{Main Takeaway: To err is human, so we must develop practices to protect ourselves from ourselves}
% \begin{itemize}
% 	\item Understand the futility of ``just trying to be careful''
% 	\item Compose code that is less likely to contains errors, and where errors that do occur are more likely to be caught.
% \end{itemize}

\end{document}
