% So we make this "beamer" rather than document!

\documentclass[11pt]{beamer}
% For handout add ,handout after 11pt

\usetheme[sectionpage=none,numbering=none]{metropolis}           % Use metropolis theme
	% To do printouts, add ", handout"  after aspectratio.
\usepackage{booktabs}
\usepackage{graphicx}
\usepackage{color}

\title{Practical Data Science I}
\author{\small Nick Eubank}
\date{\vspace*{.3in} \date}


% This is the beginning of a real document!
\begin{document}


\begin{frame}[c]
\centering
\pause THANK YOU!
\end{frame}


\begin{frame}[c]{}
You've...
  \begin{itemize}
    \item Learned to think computationally and solve problems using the 7 step method.
    \pause \item Learned to generalize problem solving logic.
    \pause \item Learned to debug programs.
    \pause \item Learned to define classes and understand object-oriented code.
    \pause \item Learned to work with the command line.
  \end{itemize}
\end{frame}

\begin{frame}[c]{}
You've...
  \begin{itemize}
   \item Measured how diplomas affect earnings (won a Nobel prize!)
    \pause \item Measured racial wage gap, racial and gender education gap, and shown how they are changing over time.
    \pause \item Documented the relationship between national income and political institutions.
    \pause \item Estimated effect of drug legalization on violent crime.
    \pause \item Estimated the causal effect of opioid regulatory policies.
  \end{itemize}
\end{frame}

\begin{frame}[c]{}
Most important insights often come from
\begin{itemize}
  \pause \item Finding and combining new sources of data, and
  \pause \item Measuring things that haven't been measured before.
\end{itemize}
\end{frame}

\begin{frame}[c]{}
\begin{center}
  You have the skills to bring evidence \\
  to bear \alert{on the questions that matter to you.}
\end{center}

\begin{itemize}
  \pause \item Work with dirty, real-world data,
  \pause \item On your own computer,
  \pause \item Even if the data is big.
  \pause \item You can also move your workflow to the cloud,
  \pause \item You have the tools to collaborate with others, get help, and contribute in the data science ecosystem.
\end{itemize}
\vspace*{0.2cm}
\pause \alert{None} of these abilities depend on faculty setting up an environment for you.\\
\end{frame}

\begin{frame}[c]{}
\centering
     \alert{Proud} and \alert{empowered}. 
  \end{frame}

\begin{frame}[c]{Evaluations}
\begin{itemize}
  \pause \item \textbf{Important for us} \\
  \pause Both for improving this class, and for informing our other classes.
  \pause \item \textbf{Important to MIDS} \\
  Data $\rightarrow$ improve our teaching!
\end{itemize}
\pause So please give these evaluations your attention for the next few minutes! \\
\pause Free-response boxes are \emph{especially} helpful!
\end{frame}

\end{document}
